---
title: "Average bereavement effect"
author: "Mathilde Marie Brünnich Sloth"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

# Average bereavement effect

The provided code is for the analysis of the average effect of bereavement on different causes of death within 3 years of index date.

```{r}
library(riskRegression)
library(dplyr)
library(survival)
library(Publish)
library(rms)
library(rsample)
library(lubridate)
library(xlsx)
```

#### Load df

Loading data frame. Code for preprocessing is also made available.

```{r}
df <- read.csv(".../Data/dataframe1.csv")
```

Formating

```{r}
df$children_n.factor <- as.factor(df$children_n.factor)
df$comorbidities_factor <- as.factor(df$comorbidities_factor)
df$immigration_status <- as.factor(df$immigration_status)
df$bereaved01 <- as.factor(df$bereaved01)
df$sex <- as.factor(df$sex)


df$surv_cancer <- as.factor(df$surv_cancer)
df$surv_cardiovascular <- as.factor(df$surv_cardiovascular)
df$surv_neuro <- as.factor(df$surv_neuro)
df$surv_dm <- as.factor(df$surv_dm)
df$surv_digestive <- as.factor(df$surv_digestive)
df$surv_psychiatric <- as.factor(df$surv_psychiatric)
df$surv_resp <- as.factor(df$surv_resp)
```

Check for missings in data.

```{r}
sum(is.na(df$children_n.factor))
sum(is.na(df$comorbidities_factor))
sum(is.na(df$immigration_status))
sum(is.na(df$average_homecare))
sum(is.na(df$average_primarycare))
sum(is.na(df$average_residential))
sum(is.na(df$average_outpatient))
sum(is.na(df$average_inpatient))
sum(is.na(df$sex))
sum(is.na(df$bereaved01))
```

## Average bereavement effect analysis

First part of the analysis males and females are analysed together.

#### Distribution of covariates

```{r}
summary(utable(bereaved01 ~ sex + age_bereavement.factor + immigration_status + comorbidities_factor + affluence.factor + children_n.factor +average_full, 
               data = df), show.totals = T)
```

#### 

#### ABE, Cancer

-   *ate* function is the average treatment effect. The first term is the formula from the CSC model.

```{r}
csc_bereavement_cancer <- CSC(Hist(surv_time, surv_cancer) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full, data = df, cause = "1")

arv_bereavement_cancer <- ate(csc_bereavement_cancer, treatment = "bereaved01", data = df, cause = "1", times = c(365*3) )

ABE_cancer <- summary(arv_bereavement_cancer,short = T,type = "ratio")

```

#### ABE, cardio and vascular disease

```{r}
csc_bereavement_cardiovascular <- CSC(Hist(surv_time, surv_cardiovascular) ~ bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df, cause = "1")

arv_bereavement_cardiovascular <- ate(csc_bereavement_cardiovascular, treatment = "bereaved01", data = df, cause = "1", times = c(365*3) )

ABE_cv <-summary(arv_bereavement_cardiovascular,short = T,type = "ratio")
```

#### ABE, Dementia and Parkinsons

```{r}
csc_bereavement_neuro <- CSC(Hist(surv_time, surv_neuro) ~ bereaved01 +age_bereavement.factor + children_n.factor + average_full, data = df, cause = "1")

arv_bereavement_neuro <- ate(csc_bereavement_neuro, treatment = "bereaved01", data = df, cause = "1", times = c(365*3) )

ABE_neuro <-summary(arv_bereavement_neuro,short = T,type = "ratio")
```

#### ABE, Diabetes

```{r}
csc_bereavement_dm <- CSC(Hist(surv_time, surv_dm) ~ bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df, cause = "1")

arv_bereavement_dm <- ate(csc_bereavement_dm, treatment = "bereaved01", data = df, cause = "1", times = c(365*3) )

ABE_dm <- summary(arv_bereavement_dm,short = T,type = "ratio")
```

#### ABE, Diseases related to the digestive system

```{r}
csc_bereavement_digestive <- CSC(Hist(surv_time, surv_digestive) ~ bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df, cause = "1")

arv_bereavement_digestive <- ate(csc_bereavement_digestive, treatment = "bereaved01", data = df, cause = "1", times = c(365*3) )

ABE_digestive <- summary(arv_bereavement_digestive,short = T,type = "ratio")
```

#### ABE, Psychiatric diseases and suicide

```{r}
csc_bereavement_psychiatric <- CSC(Hist(surv_time, surv_psychiatric) ~ bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df, cause = "1")

arv_bereavement_psychiatric <- ate(csc_bereavement_psychiatric, treatment = "bereaved01", data = df, cause = "1", times = c(365, 365*2, 365*3) )

ABE_psychiatric <- summary(arv_bereavement_psychiatric,short = T,type = "ratio")
```

#### ABE, Respiratory diseases

```{r}
csc_bereavement_resp <- CSC(Hist(surv_time, surv_resp) ~ bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df, cause = "1")

arv_bereavement_resp <- ate(csc_bereavement_resp, treatment = "bereaved01", data = df, cause = "1", times = c( 365*3) )

ABE_resp <- summary(arv_bereavement_resp,short = T,type = "ratio")

```

#### ABE, together

```{r}
df_ABE_cancer <- ABE_cancer$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Cancer")

df_ABE_cv <- ABE_cv$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "CV")
  
df_ABE_neuro <- ABE_neuro$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Neuro")

df_ABE_dm <- ABE_dm$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Diabetes")

df_ABE_digestive <- ABE_digestive$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Digestive")

  
df_ABE_psychiatric <- ABE_psychiatric$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Psychiatric")

df_ABE_resp <- ABE_resp$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Resp")


df_ABE_together <- df_ABE_cancer |> 
  full_join(df_ABE_cv) |> 
  full_join(df_ABE_neuro) |> 
  full_join(df_ABE_dm) |> 
  full_join(df_ABE_digestive) |> 
  full_join(df_ABE_psychiatric) |> 
  full_join(df_ABE_resp) |> 
  select(ABE, time, A, estimate.A, B, estimate.B, estimate, lower, upper, p.value)



write.xlsx(df_ABE_together, ".../Results/ABE.xlsx")


```

## ABE, Sex differences

The average bereavement effect is analysed separately for males and females.

### Split data by sex

```{r}

df_males <- df |> filter(sex == "1")

df_females <- df |> filter(sex == "2")

table_males <- summary(utable(bereaved01 ~ age_bereavement.factor + immigration_status + comorbidities_factor + affluence.factor + children_n.factor +average_primarycare + average_inpatient + average_outpatient + average_residential + average_homecare + average_prescription, 
               data = df_males), show.totals = T)


table_females <- summary(utable(bereaved01 ~ age_bereavement.factor + immigration_status + comorbidities_factor + affluence.factor + children_n.factor +average_primarycare + average_inpatient + average_outpatient + average_residential + average_homecare + average_prescription, 
               data = df_females), show.totals = T)

```

#### ABE, Cancer

-   *ate* function is the average treatment effect. The first term is the formula from the CSC model.

```{r}

# Males

df_males$bereaved01 <- as.factor(df_males$bereaved01)

csc_bereavement_cancer_males <- CSC(Hist(surv_time, surv_cancer) ~ bereaved01 + average_full, data = df_males, cause = "1")

arv_bereavement_cancer_males <- ate(csc_bereavement_cancer_males, treatment = "bereaved01", data = df_males, cause = "1", times = c(365*3))

ABE_cancer_males <- summary(arv_bereavement_cancer_males,short = T,type = "ratio")


# Females


df_females$bereaved01 <- as.factor(df_females$bereaved01)

csc_bereavement_cancer_females <- CSC(Hist(surv_time, surv_cancer) ~ bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df_females, cause = "1")

arv_bereavement_cancer_females <- ate(csc_bereavement_cancer_females, treatment = "bereaved01", data = df_females, cause = "1", times = c(365*3))

ABE_cancer_females <- summary(arv_bereavement_cancer_females,short = T,type = "ratio")
```

#### ABE, cardio and vascular disease

```{r}

# Males

csc_bereavement_cardiovascular_males <- CSC(Hist(surv_time, surv_cardiovascular) ~ bereaved01 + average_full, data = df_males, cause = "1")

arv_bereavement_cardiovascular_males <- ate(csc_bereavement_cardiovascular_males, treatment = "bereaved01", data = df_males, cause = "1", times = c(365*3))

ABE_cv_males <- summary(arv_bereavement_cardiovascular_males,short = T,type = "ratio")


# Females

csc_bereavement_cardiovascular_females <- CSC(Hist(surv_time, surv_cardiovascular) ~  bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df_females, cause = "1")

arv_bereavement_cardiovascular_females <- ate(csc_bereavement_cardiovascular_females, treatment = "bereaved01", data = df_females, cause = "1", times = c(365*3))

ABE_cardiovascular_females <- summary(arv_bereavement_cardiovascular_females,short = T,type = "ratio")
```

#### ABE, Dementia and Parkinsons

```{r}

# Males

csc_bereavement_neuro_males <- CSC(Hist(surv_time, surv_neuro) ~ bereaved01 + average_full, data = df_males, cause = "1")

arv_bereavement_neuro_males <- ate(csc_bereavement_neuro_males, treatment = "bereaved01", data = df_males, cause = "1", times = c(365*3))

ABE_neuro_males <-summary(arv_bereavement_neuro_males,short = T,type = "ratio")


# Females

csc_bereavement_neuro_females <- CSC(Hist(surv_time, surv_neuro) ~  bereaved01 + age_bereavement.factor + children_n.factor + average_full , data = df_females, cause = "1")

arv_bereavement_neuro_females <- ate(csc_bereavement_neuro_females, treatment = "bereaved01", data = df_females, cause = "1", times = c(365*3))

ABE_neuro_females <- summary(arv_bereavement_neuro_females,short = T,type = "ratio")
```

#### ABE, Diabetes

```{r}

# Males

csc_bereavement_dm_males <- CSC(Hist(surv_time, surv_dm) ~ bereaved01 + average_full, data = df_males, cause = "1")

arv_bereavement_dm_males <- ate(csc_bereavement_dm_males, treatment = "bereaved01", data = df_males, cause = "1", times = c(365*3))

ABE_dm_males <- summary(arv_bereavement_dm_males,short = T,type = "ratio")


# Females

csc_bereavement_dm_females <- CSC(Hist(surv_time, surv_dm) ~  bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df_females, cause = "1")

arv_bereavement_dm_females <- ate(csc_bereavement_dm_females, treatment = "bereaved01", data = df_females, cause = "1", times = c(365*3))

ABE_dm_females <- summary(arv_bereavement_dm_females,short = T,type = "ratio")
```

#### ABE, Diseases related to the digestive system

```{r}

# Males

csc_bereavement_digestive_males <- CSC(Hist(surv_time, surv_digestive) ~ bereaved01 + average_full, data = df_males, cause = "1")

arv_bereavement_digestive_males <- ate(csc_bereavement_digestive_males, treatment = "bereaved01", data = df_males, cause = "1", times = c(365*3))

ABE_digestive_males <- summary(arv_bereavement_digestive_males,short = T,type = "ratio")


# Females

csc_bereavement_digestive_females <- CSC(Hist(surv_time, surv_digestive) ~  bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df_females, cause = "1")

arv_bereavement_digestive_females <- ate(csc_bereavement_digestive_females, treatment = "bereaved01", data = df_females, cause = "1", times = c(365*3))

ABE_digestive_females <- summary(arv_bereavement_digestive_females,short = T,type = "ratio")
```

#### ABE, Psychiatric diseases and suicide

```{r}

# Males

csc_bereavement_psychiatric_males <- CSC(Hist(surv_time, surv_psychiatric) ~ bereaved01 + average_full, data = df_males, cause = "1")

arv_bereavement_psychiatric_males <- ate(csc_bereavement_psychiatric_males, treatment = "bereaved01", data = df_males, cause = "1", times = 365*3)

ABE_psychiatric_males <- summary(arv_bereavement_psychiatric_males,short = T,type = "ratio")

# Females

csc_bereavement_psychiatric_females <- CSC(Hist(surv_time, surv_psychiatric) ~  bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df_females, cause = "1")

arv_bereavement_psychiatric_females <- ate(csc_bereavement_psychiatric_females, treatment = "bereaved01", data = df_females, cause = "1", times = 365*3)

ABE_psychiatric_females <- summary(arv_bereavement_psychiatric_females,short = T,type = "diffRisk")
```

#### ABE, Respiratory diseases

```{r}

# Males

csc_bereavement_resp_males <- CSC(Hist(surv_time, surv_resp) ~  bereaved01 + average_full, data = df_males, cause = "1")

arv_bereavement_resp_males <- ate(csc_bereavement_resp_males, treatment = "bereaved01", data = df_males, cause = "1", times = c( 365*3))

ABE_resp_males <- summary(arv_bereavement_resp_males,short = T,type = "ratio")


# Females

csc_bereavement_resp_females <- CSC(Hist(surv_time, surv_resp) ~  bereaved01 + age_bereavement.factor + children_n.factor + average_full, data = df_females, cause = "1")

arv_bereavement_resp_females <- ate(csc_bereavement_resp_females, treatment = "bereaved01", data = df_females, cause = "1", times = c( 365*3))

ABE_resp_females <- summary(arv_bereavement_resp_females,short = T,type = "ratio")
```

#### ABE, together

```{r}

#Males

df_ABE_cancer_males <- ABE_cancer_males$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Cancer")

df_ABE_cv_males <- ABE_cv_males$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "CV")
  
df_ABE_neuro_males <- ABE_neuro_males$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Neuro")

df_ABE_dm_males <- ABE_dm_males$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Diabetes")

df_ABE_digestive_males <- ABE_digestive_males$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Digestive")

  
df_ABE_psychiatric_males <- ABE_psychiatric_males$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Psychiatric")

df_ABE_resp_males <- ABE_resp_males$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Resp")


df_ABE_together_males <- df_ABE_cancer_males |> 
  full_join(df_ABE_cv_males) |> 
  full_join(df_ABE_neuro_males) |> 
  full_join(df_ABE_dm_males) |> 
  full_join(df_ABE_digestive_males) |> 
  full_join(df_ABE_psychiatric_males) |> 
  full_join(df_ABE_resp_males) |> 
  select(ABE, A, estimate.A, B, estimate.B, estimate, lower, upper, p.value)



write.xlsx(df_ABE_together_males, ".../ABE_male.xlsx")


# Females

df_ABE_cancer_females <- ABE_cancer_females$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Cancer")

df_ABE_cv_females <- ABE_cardiovascular_females$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "CV")
  
df_ABE_neuro_females <- ABE_neuro_females$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Neuro")

df_ABE_dm_females <- ABE_dm_females$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Diabetes")

df_ABE_digestive_females <- ABE_digestive_females$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Digestive")

df_ABE_psychiatric_females <- ABE_psychiatric_females$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Psychiatric")

df_ABE_resp_females <- ABE_resp_females$ratioRisk |> 
  as.data.frame() |> 
  mutate(ABE = "Resp")


df_ABE_together_females <- df_ABE_cancer_females |> 
  full_join(df_ABE_cv_females) |> 
  full_join(df_ABE_neuro_females) |> 
  full_join(df_ABE_dm_females) |> 
  full_join(df_ABE_digestive_females) |> 
  full_join(df_ABE_psychiatric_females) |> 
  full_join(df_ABE_resp_females) |> 
  select(ABE, time, A, estimate.A, B, estimate.B, estimate, lower, upper, p.value)


write.xlsx(df_ABE_together_females, ".../ABE_female.xlsx")


```
