---
title: "Predicting cause-specific mortality in older Danish adults"
author: "Mathilde Marie Brünnich Sloth"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

Load libraries

```{r}
library(riskRegression)
library(dplyr)
library(survival)
library(Publish)
library(rms)
library(rsample)
library(lubridate)
library(tidymodels)
library(xlsx)
library(finetune) #tuning hyperparameters
library(DALEX) #explainability
library(DALEXtra) #explainability
library(ingredients) #explainability
library(ggsci) #colours for plots
library(patchwork) #arrangeing plots together
library(extrafont) #fonts for ggplot
```

Load data

```{r}
df <- read.csv(".../Data/dataframe1.csv")

setwd("...")
```

Format

```{r}
df$children_n.factor <- as.factor(df$children_n.factor)
df$comorbidities_factor <- as.factor(df$comorbidities_factor)
df$immigration_status <- as.factor(df$immigration_status)
df$bereaved01 <- as.factor(df$bereaved01)

df$surv_cancer <- as.factor(df$surv_cancer)
df$surv_cardiovascular <- as.factor(df$surv_cardiovascular)
df$surv_neuro <- as.factor(df$surv_neuro)
df$surv_dm <- as.factor(df$surv_dm)
df$surv_digestive <- as.factor(df$surv_digestive)
df$surv_psychiatric <- as.factor(df$surv_psychiatric)
df$surv_resp <- as.factor(df$surv_resp)

df <- df |> 
  mutate_if(is.character,as.factor)

df <- df |> 
  select(-EXCLUDE, -EXCLUDE_REASON)
```

## Binary outcome

Death due to specific causes is remade into a binary framework with 0 being alive or dying from another cause and 1 being dying from the cause in focus.

```{r}
df <- df |> 
  mutate(surv_cancer01 = if_else(surv_cancer==0 | surv_cancer==2, 0, 1)) |> 
  mutate(surv_CVD01 = if_else(surv_cardiovascular==0 | surv_cardiovascular==2, 0, 1)) |> 
  mutate(surv_neuro01 = if_else(surv_neuro==0 | surv_neuro==2, 0, 1)) |> 
  mutate(surv_dm01 = if_else(surv_dm==0 | surv_dm==2, 0, 1)) |> 
  mutate(surv_digestive01 = if_else(surv_digestive==0 | surv_digestive==2, 0, 1)) |> 
  mutate(surv_psychiatric01 = if_else(surv_psychiatric==0 | surv_psychiatric==2, 0, 1)) |> 
  mutate(surv_respiratory01 = if_else(surv_resp==0 | surv_resp==2, 0, 1))
```

## Survival (1 year and 3 years)

Computing a binary variable for 1 and 3-year survival for each cause of death.

```{r}
df <- df |> 
  mutate(surv_cancer3 = if_else(surv_cancer == 1 & surv_time <= 1095.75, 1,0),
         surv_CVD3 =  if_else(surv_cardiovascular == 1 & surv_time <= 1095.75, 1,0),
         surv_neuro3 =  if_else(surv_neuro == 1 & surv_time <= 1095.75, 1,0),
         surv_dm3 =  if_else(surv_dm == 1 & surv_time <= 1095.75, 1,0),
         surv_digestive3 =  if_else(surv_digestive == 1 & surv_time <= 1095.75, 1,0),
         surv_psychiatric3 =  if_else(surv_psychiatric == 1 & surv_time <= 1095.75, 1,0),
         surv_psychiatric2 =  if_else(surv_psychiatric == 1 & surv_time <= 730.5, 1,0),
         surv_psychiatric1 =  if_else(surv_psychiatric == 1 & surv_time <= 365.25, 1,0),
         surv_respiratory3 =  if_else(surv_resp == 1 & surv_time <= 1095.75, 1,0),
         surv_overall3 =  if_else(surv_overall == 1 & surv_time <= 1095.75, 1,0)) |> 
  mutate(surv_cancer3 = as.factor(surv_cancer3),
         surv_CVD3 = as.factor(surv_CVD3),
         surv_neuro3 = as.factor(surv_neuro3),
         surv_dm3 = as.factor(surv_dm3),
         surv_digestive3 = as.factor(surv_digestive3),
         surv_psychiatric3 = as.factor(surv_psychiatric3),
         surv_psychiatric2 = as.factor(surv_psychiatric2),
         surv_psychiatric1 = as.factor(surv_psychiatric1),
         surv_respiratory3 = as.factor(surv_respiratory3) )


df |> janitor::tabyl(surv_cancer3)
```

## Levels

Naming the categories of the survival variable.

```{r}
levels(df$surv_overall)[levels(df$surv_overall)== 0] <- 'Alive'
levels(df$surv_overall)[levels(df$surv_overall)== 1] <- 'Dead'
```

## Prediction

First step of the prediction modelling is a data split into train/test.

```{r}
set.seed(29)

df_split <- initial_split(df)

df_train <- training(df_split)

df_test <- testing(df_split)
```

Descriptive statistics

```{r}

df_train$TRAIN <- "TRUE"
df_test$TEST <- "TRUE"

df_look <- df_train |>
  full_join(df_test) |> 
  mutate(TRAIN_TEST = case_when(TRAIN == "TRUE" ~ "TRAIN", 
                                TEST == "TRUE" ~ "TEST"))

#look at the distribution of the outcome with the utable function
death_cause_tabel <- utable(TRAIN_TEST ~ surv_cancer3 + surv_CVD3 + surv_neuro3 + surv_dm3 + surv_digestive3 + surv_psychiatric3 + surv_respiratory3 + surv_overall3, data = df_look) 

#Cause of death table
death_cause_tabel <- summary(death_cause_tabel)
write.xlsx(death_cause_tabel, ".../Results/Table_cause_of_death.xls")

#same for males and females sep
death_cause_tabel_sex <- utable(sex ~ surv_cancer3 + surv_CVD3 + surv_neuro3 + surv_dm3 + surv_digestive3 + surv_psychiatric3 + surv_respiratory3 , data = df_look) 
death_cause_tabel_sex_table <- summary(death_cause_tabel_sex)
write.xlsx(death_cause_tabel_sex_table, ".../Results/Table_cause_of_death_sex.xls")

#table 1
table_1 <- summary(utable(TRAIN_TEST ~ bereaved01 + sex + age_bereavement.factor + immigration_status + comorbidities_factor + children_n.factor + affluence.factor + average_full + average_prescription + average_homecare + average_residential + average_primarycare + average_outpatient + average_inpatient, 
               data = df_look, show.totals = T, summary.format = "median(x) [iqr(x)]"))

write.xlsx(table_1, ".../Results/Table1.xls")


subset_test <- subset(df_look, TRAIN_TEST= "TEST" )
subset_train <- subset(df_look, TRAIN_TEST= "TRAIN" )

#prescription
quantile(df_look$average_prescription, prob=c(.05,.95))
quantile(subset_train$average_prescription, prob=c(.05,.95))
quantile(subset_test$average_prescription, prob=c(.05,.95)) 

#homecare
quantile(df_look$average_homecare, prob=c(.05,.95))
quantile(subset_train$average_homecare, prob=c(.05,.95))
quantile(subset_test$average_homecare, prob=c(.05,.95)) 

#residential care
quantile(df_look$average_residential, prob=c(.05,.95))
quantile(subset_train$average_residential, prob=c(.05,.95))
quantile(subset_test$average_residential, prob=c(.05,.95)) 

#primary care
quantile(df_look$average_primarycare, prob=c(.05,.95))
quantile(subset_train$average_primarycare, prob=c(.05,.95))
quantile(subset_test$average_primarycare, prob=c(.05,.95)) 

#outpatient
quantile(df_look$average_outpatient, prob=c(.05,.95))
quantile(subset_train$average_outpatient, prob=c(.05,.95))
quantile(subset_test$average_outpatient, prob=c(.05,.95)) 

#inpatient
quantile(df_look$average_inpatient, prob=c(.05,.95))
quantile(subset_train$average_inpatient, prob=c(.05,.95))
quantile(subset_test$average_inpatient, prob=c(.05,.95)) 

#medians with more decimal points
median(df_look$average_prescription)
median(subset_train$average_prescription)
median(subset_test$average_prescription)
# 0.04

median(df_look$average_homecare)
#0

median(df_look$average_residential)
#0

median(df_look$average_primarycare)
#0.019

median(df_look$average_outpatient)
#0.036

median(df_look$average_inpatient)
#0

```

### Logistic regression

First model for the prediction modelling is logistic regression models. Three prediction models for each cause of death will be performed:

-   LRM null: a logistic regression model (LRM) without any predictors

-   LRM simple: a LRM with the predictors age, sex and bereavement

-   LRM full: a LRM with all predictors: age, sex, bereavement, affluence, immigration status, number of children, number of comorbidities, and expenditures with restricted cubic splines with 3 knots.

Investigating the predicted risks for 3 year mortality of cause-specific death.

Specify data distance

```{r}
dd <- datadist(df_train)

options(datadist = "dd")
```

##### All cause mortality

```{r}
fit_overall_null <- glm(surv_overall3 ~ 1, data = df_train, x = T, y = T, family = "binomial")

fit_overall <- lrm(surv_overall3 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor * rcs(average_full,3) + children_n.factor + comorbidities_factor + immigration_status, data = df_train, x = T, y = T)

fit1_score <- Score(object = list("LRM Full Model, overall" = fit_overall, 
                                  "Null Model" = fit_overall_null), 
                    data = df_test,
                    formula = surv_overall3 ~ 1,
                    metrics = c("auc","brier"), summary = "ipa", plots = "cal",
                    se.fit = T,  null.model = T) 
                    
plotCalibration(fit1_score,legend = F)
```

##### Cancer

```{r}
#Null model
fit0_cancer <- glm(surv_cancer3 ~ 1, data = df_train, family = "binomial",x = T, y = T)

#simple model with age and sex
fit_cancer_simple <- lrm(surv_cancer3 ~ bereaved01 + sex + age_bereavement.factor, data = df_train, x = T, y = T)

#model with expenditures as arv.
fit_cancer_arv <- lrm(surv_cancer3 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor + children_n.factor + comorbidities_factor + immigration_status + rcs(average_inpatient, 3) +rcs(average_outpatient, 3) +  rcs(average_prescription, 3) + rcs(average_primarycare, 3) + rcs(average_residential,3) + rcs(average_homecare,3), 
            data = df_train, x = T, y = T)
```

Evaluating the scores

```{r}
df_test <- data.table::as.data.table(df_test)

fit_cancer_total_score <- Score(object = 
                          list("LRM Simple Model" = fit_cancer_simple,
                                "LRM Full Model, each arv exp" = fit_cancer_arv,
                                "LRM Null" = fit0_cancer),
                    data = df_test,
                    formula = surv_cancer3 ~ 1,
                    metrics = c("auc","brier"),summary = "ipa",plots = "cal",
                    se.fit = T,  null.model = T) 
  
summary(fit_cancer_total_score)

plotCalibration(fit_cancer_total_score, models = "LRM Full Model, each arv exp")
```

Nomogram

```{r}
uu <- datadist(df)
options(datadist="uu")
plot(nomogram(fit_cancer_arv, fun = plogis, funlabel = paste0("Risk of cancer")))
```

##### Cardio and vascular diseases

```{r}
#Null model
fit0_cv <- glm(surv_CVD3 ~ 1, data = df_train, family = "binomial", x = T, y = T)

#simple model with age and sex
fit_cv_simple <- lrm(surv_CVD3 ~ bereaved01 + sex + age_bereavement.factor, data = df_train, x = T, y = T)

#model with expenditures as arv.
fit_cv_arv <- lrm(surv_CVD3 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor + children_n.factor + comorbidities_factor + immigration_status + rcs(average_inpatient, 3) +rcs(average_outpatient, 3) +  rcs(average_prescription, 3) + rcs(average_primarycare, 3) + rcs(average_residential,3) + rcs(average_homecare,3), 
            data = df_train, x = T, y = T)
```

Evaluating the scores

```{r}
fit_cv_total_score <- Score(object = 
                            list("LRM Simple Model" = fit_cv_simple,
                                 "LRM Full Model, each arv exp" = fit_cv_arv,
                                 "LRM Null" = fit0_cv),
                    data = df_test,
                    formula = surv_CVD3 ~ 1,
                    metrics = c("auc","brier"), summary = "ipa", plots = "cal",
                    se.fit = T, null.model = T)

summary(fit_cv_total_score)

plotCalibration(fit_cv_total_score, models = "LRM Full Model, each arv exp")
plotCalibration(fit_cv_total_score, models = "LRM Simple Model")
```

Nomogram

```{r}
uu <- datadist(df)
options(datadist="uu")
plot(nomogram(fit_cv_arv, fun = plogis, funlabel = paste0("Risk of CVD")))
```

##### Dementia and Parkinson disease

```{r}
#Null model
fit0_neuro <- glm(surv_neuro3 ~ 1, data = df_train, family = "binomial", x = T, y = T)

#simple model with age and sex
fit_neuro_simple <- lrm(surv_neuro3 ~ bereaved01 + sex + age_bereavement.factor, data = df_train, x = T, y = T)

#model with expenditures as arv.
fit_neuro_arv <- lrm(surv_neuro3 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor + children_n.factor + comorbidities_factor + immigration_status + rcs(average_inpatient, 3) +rcs(average_outpatient, 3) +  rcs(average_prescription, 3) + rcs(average_primarycare, 3) + rcs(average_residential,3) + rcs(average_homecare,3), 
            data = df_train, x = T, y = T)
```

Evaluate the scores

```{r}
fit_neuro_total_score <- Score(object = 
                        list("LRM Simple Model" = fit_neuro_simple,
                             "LRM Full Model, each arv exp" = fit_neuro_arv,
                             "LRM Null" = fit0_neuro),
                    data = df_test,
                    formula = surv_neuro3 ~ 1,
                    metrics = c("auc","brier"), summary = "ipa",plots = "cal",
                    se.fit = T, 
                     null.model = T)

summary(fit_neuro_total_score)

plotCalibration(fit_neuro_total_score)
```

Nomogram

```{r}
uu <- datadist(df)
options(datadist="uu")
plot(nomogram(fit_neuro_simple, fun = plogis, funlabel = paste0("Risk of dementia or parkinsons")))
plot(nomogram(fit_neuro_arv, fun = plogis, funlabel = paste0("Risk of dementia or parkinsons")))
```

##### Diabetes

```{r}
#Null model
fit0_dm <- glm(surv_dm3 ~ 1, data = df_train, family = "binomial")

#simple model with age and sex
fit_dm_simple <- lrm(surv_dm3 ~ bereaved01 + sex + age_bereavement.factor, data = df_train, x = T, y = T)

#model with expenditures as arv.
fit_dm_arv <- lrm(surv_dm3 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor + children_n.factor + comorbidities_factor + immigration_status + rcs(average_inpatient, 3) +rcs(average_outpatient, 3) +  rcs(average_prescription, 3) + rcs(average_primarycare, 3) + rcs(average_residential,3) + rcs(average_homecare,3), 
            data = df_train, x = T, y = T)
```

Evaluate score

```{r}
fit_dm_total_score <- Score(object = 
                            list("LRM Simple Model" = fit_dm_simple,
                                 "LRM Full Model, each arv exp" = fit_dm_arv,
                                 "LRM Null" = fit0_dm), 
                    data = df_test,
                    formula = surv_dm3 ~ 1,
                    metrics = c("auc","brier"),summary = "ipa",plots = "cal",
                    se.fit = T,
                    null.model = T)

summary(fit_dm_total_score)

plotCalibration(fit_dm_total_score, models = "LRM Full Model, each arv exp",round = F)

plot(nomogram(fit_dm_arv, fun = plogis, funlabel = paste0("Risk of Diabetes")))
```

##### Diseases related to the digestive system

```{r}
#Null model
fit0_digestive <- glm(surv_digestive3 ~ 1, data = df_train, family = "binomial", x = T, y = T)

#simple model with age and sex
fit_digestive_simple <- lrm(surv_digestive3 ~ bereaved01 + sex + age_bereavement.factor, data = df_train, x = T, y = T)

#model with expenditures as arv.
fit_digestive_arv <- lrm(surv_digestive3 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor + children_n.factor + comorbidities_factor + immigration_status + rcs(average_inpatient, 3) +rcs(average_outpatient, 3) +  rcs(average_prescription, 3) + rcs(average_primarycare, 3) + rcs(average_residential,3) + rcs(average_homecare,3), 
            data = df_train, x = T, y = T)
```

Evaluate score

```{r}
fit_digestive_total_score <- Score(object = 
                            list("LRM Simple Model" = fit_digestive_simple,
                                 "LRM Full Model, each arv exp" = fit_digestive_arv,
                                 "LRM Null" = fit0_digestive),
                    data = df_test,
                    formula = surv_digestive3 ~ 1,
                    metrics = c("auc","brier"),summary = "ipa",plots = "cal",
                    se.fit = T,
                    null.model = T)

summary(fit_digestive_total_score)

plotCalibration(fit_digestive_total_score, models = "LRM Full Model, each arv exp", round = F)

plot(nomogram(fit_digestive_arv, fun = plogis, funlabel = paste0("Risk of digestive diseases")))
```

##### Psychiatric diseases and suicide

```{r}
#Null model
fit0_psychiatric <- glm(surv_psychiatric3 ~ 1, data = df_train, family = "binomial", x = T, y = T)

#simple model with age and sex
fit_psychiatric_simple <- lrm(surv_psychiatric3 ~ bereaved01 + sex + age_bereavement.factor, data = df_train, x = T, y = T)

#model with expenditures as arv.
fit_psychiatric_arv <- lrm(surv_psychiatric3 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor + children_n.factor + comorbidities_factor + immigration_status + rcs(average_inpatient, 3) +rcs(average_outpatient, 3) +  rcs(average_prescription, 3) + rcs(average_primarycare, 3) + rcs(average_residential,3) + rcs(average_homecare,3), 
            data = df_train, x = T, y = T)
```

Evaluate score

```{r}
fit_psychiatric_total_score <- Score(object = 
                            list("LRM Simple Model" = fit_psychiatric_simple,
                                 "LRM Full Model, each arv exp" = fit_psychiatric_arv,
                                 "LRM Null" = fit0_psychiatric),
                    data = df_test,
                    formula = surv_psychiatric3 ~ 1,
                    metrics = c("auc","brier"), summary = "ipa", plots = "cal",
                    se.fit = T,  null.model = T)

summary(fit_psychiatric_total_score)

plotCalibration(fit_psychiatric_total_score, models = "LRM Full Model, each arv exp", round = F)

plot(nomogram(fit_psychiatric_arv, fun = plogis, funlabel = paste0("Risk of psychiatric diseases")))
```

1 year psych (not included in thesis)

```{r}
#Null model
fit0_psychiatric1 <- glm(surv_psychiatric1 ~ 1, data = df_train, family = "binomial", x = T, y = T)

#simple model with age and sex
fit_psychiatric_simple1 <- lrm(surv_psychiatric1 ~ bereaved01 + sex + age_bereavement.factor, data = df_train, x = T, y = T)

#model with expenditures as arv.
fit_psychiatric_arv1 <- lrm(surv_psychiatric1 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor + children_n.factor + comorbidities_factor + immigration_status + rcs(average_inpatient, 3) +rcs(average_outpatient, 3) +  rcs(average_prescription, 3) + rcs(average_primarycare, 3) + rcs(average_residential,3) + rcs(average_homecare,3), 
            data = df_train, x = T, y = T)

fit_psychiatric_total_score1 <- Score(object = 
                            list("LRM Simple Model 1y" = fit_psychiatric_simple1,
                                 "LRM Full Model, each arv exp 1y" = fit_psychiatric_arv1,
                                 "LRM Null 1y" = fit0_psychiatric1),
                    data = df_test,
                    formula = surv_psychiatric1 ~ 1,
                    metrics = c("auc","brier"), summary = "ipa", plots = "cal",
                    se.fit = T,  null.model = T)
```

##### Respiratory diseases

```{r}
#Null model
fit0_resp <- glm(surv_respiratory3 ~ 1, data = df_train, family = "binomial", x = T, y = T)

#simple model with age and sex
fit_resp_simple <- lrm(surv_respiratory3 ~ bereaved01 + sex + age_bereavement.factor, data = df_train, x = T, y = T)

#model with expenditures as arv.
fit_resp_arv <- lrm(surv_respiratory3 ~ bereaved01 + affluence.factor + sex + age_bereavement.factor + children_n.factor + comorbidities_factor + immigration_status + rcs(average_inpatient, 3) +rcs(average_outpatient, 3) +  rcs(average_prescription, 3) + rcs(average_primarycare, 3) + rcs(average_residential,3) + rcs(average_homecare,3), 
            data = df_train, x = T, y = T)
```

Evaluate score

```{r}
fit_resp_total_score <- Score(object = 
                         list("LRM Simple Model" = fit_resp_simple,
                              "LRM Full Model, each arv exp" = fit_resp_arv,
                              "LRM Null" = fit0_resp),
                    data = df_test,
                    formula = surv_respiratory3 ~ 1,
                    metrics = c("auc","brier"),summary = "ipa", plots = "cal",
                    se.fit = T, 
                    null.model = T)

summary(fit_resp_total_score)

plotCalibration(fit_resp_total_score, models = "LRM Full Model, each arv exp")

plot(nomogram(fit_resp_arv, fun = plogis, funlabel = paste0("Risk of respiratory diseases")))
```

##### 

### XGBoost

Next model is Extreme Gradient Boosting (XGBoost) with all predictors for predicting the 3-year mortality.

Cross-validations split for tuning hyper-parameters

```{r}
set.seed(29)

df_train_cv <- vfold_cv(data = df_train, v = 5)
```

#### Cancer

For this analysis the outcome must be a factor. Catagorical predictors should have dummy variables.

-   1: get the disease (cancer) before end of follow-up
-   0: does not get the disease (dies of something else or survives until follow-up)

```{r}
#writing recipe for the formula and making the dummy variables
xg_cancer_recipe <- recipe(surv_cancer3 ~ bereaved01 + affluence.factor +                                            sex + age_bereavement.factor + 
                           children_n.factor + comorbidities_factor +
                           immigration_status + average_inpatient +
                           average_outpatient + average_prescription +
                           average_primarycare + average_residential +
                           average_homecare, data = df_train) |> 
  step_dummy(all_nominal_predictors(),one_hot = T)

#xg boost with computation engine: xgboost
xgb_cancer_spec <- boost_tree(trees = tune(),
                              mtry = tune(),
                              min_n = tune(),
                              tree_depth = tune(),
                              stop_iter = tune(),
                              learn_rate = tune()
                              ) |> 
                              set_engine("xgboost", validation = 0.2) |> 
                              set_mode("classification")

#workflow
xgb_cancer_wf <- workflow(xg_cancer_recipe, xgb_cancer_spec)


#tune hyperparameters (defines how our model is structured). Use the train df with crossvalidation.

set.seed(29)

doParallel::registerDoParallel()

xgb_cancer_rs <- tune_race_anova(xgb_cancer_wf,
                          df_train_cv,
                          grid = 30,
                          metrics = metric_set(brier_class),
                          control = control_race(verbose_elim = TRUE))

#plot results
plot_race(xgb_cancer_rs)

#Grid of hyperparameters
show_best(xgb_cancer_rs)


#finalize workflow
xgb_cancer_last <- xgb_cancer_wf |> 
    finalize_workflow(select_best(xgb_cancer_rs, 'brier_class')) |> 
    last_fit(df_split)

#metrics
xgb_cancer_last |> collect_metrics()

#predictions
xgb_cancer_last |> collect_predictions()


#group
xgb_cancer_full_preds <- xgb_cancer_last$.predictions

xgb_cancer_full_preds <- xgb_cancer_full_preds[[1]][2]

xgb_cancer_full_preds <- as.matrix(xgb_cancer_full_preds)




#Specify datadist
dd <- datadist(df_train)

options(datadist = "dd")

# evaluate performance
fit_cancer_total_score <- Score(list('LRM NULL' = fit0_cancer,
                              'LRM Simple' = fit_cancer_simple,
                              'LRM Full' = fit_cancer_arv,
                              'Xgboost Full' = xgb_cancer_full_preds), 
                               formula = surv_cancer3 ~ 1,
                               data = df_test,
                               summary = 'ipa', plots = 'calibration',
                               metrics = c('AUC', 'brier'), 
                               se.fit = T)
summary(fit_cancer_total_score)

plotCalibration(fit_cancer_total_score, models =  "Xgboost Full", legend = T, auc.in.legend = F, brier.in.legend = F, col = c("#88CCEE", "#117733"), rug = F, xlab = "Predicted risk of death from cancer", ylab = "Observed frequency of death from cancer")



```

##### Prediction plots

```{r}
#put the predictions from the xgb into the df. 
df_test$Predictions_Cancer <-  xgb_cancer_full_preds

#personal colour scheme
m_color <- c("#56B4E9", "#CC79A7", "#0072B2", "#E69F00", "#009E73")


#sex differenes and age
sex_age_cancer <- df_test |>
  ggplot(aes(x = age_bereavement, y = Predictions_Cancer, fill=bereaved01, colour = bereaved01)) +
  geom_smooth(show.legend = F) +
  facet_wrap(vars(sex))+
  theme_minimal(base_size = 14) +
  scale_color_manual(values = m_color)+ 
  scale_fill_manual(values = m_color)+  
  xlab("Age at bereavement") +
  ylab("Predicted risk of dying from cancer") +
  scale_x_continuous(limits = c(65, 100)) 

#save
pdf(".../Results/Cancer/Sex differences/sex_age_cancer.pdf", width = 7, height = 7)
print(sex_age_cancer)
dev.off()

#geoim_smooth(..., method = loess)

#Differences in predicted risks for bereaved and non-bereaved.
boxplot_cancer <- df_test |>
          ggplot(aes(x = bereaved01, y = Predictions_Cancer)) +
          geom_boxplot(outlier.shape = NA, color = "navy", fill = "#56B4E9", alpha=0.6) + 
          theme_minimal(base_size = 14) +
          facet_wrap(vars(sex))+
          scale_y_continuous(limits = c(0, 0.15)) + 
          xlab("Bereavement status") +
          ylab("Predicted risk of dying from cancer")


pdf(".../Results/Cancer/boxplot_cancer.pdf", width = 7, height = 7)
print(boxplot_cancer)
dev.off()

#distribution of risks
cancer_risk_distribution <- df_test |> 
              ggplot(aes(x = Predictions_Cancer,fill = surv_cancer3)) + 
              geom_density(alpha = 0.4, show.legend = F) +
              xlab("Predicted risk of dying from cancer") +
              theme_minimal(base_size = 14) +
              scale_color_manual(values = m_color)+ 
              scale_fill_manual(values = m_color)

#save
pdf(".../Results/Cancer/Sex differences/cancer_risk_distribution.pdf", width = 7, height = 7)
print(cancer_risk_distribution)
dev.off()


############ CURE RELATED EXP #################
#blue: non-bereaved

#expenditures, sex diffrences
cancer_inpatient <- df_test |>
                ggplot(aes(x = average_inpatient, y = Predictions_Cancer, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend =F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
   theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital inpatient expenditure") +
                        ylab("Predicted risk of dying from cancer") +
                        scale_x_continuous(limits = c(0, 1)) 


#test
cancer_inpatient <- df_test |>
                ggplot(aes(x = average_inpatient, y = Predictions_Cancer, 
                           fill=bereaved01, colour = bereaved01)) +
                        Hmisc::histSpikeg(Predictions_Cancer ~ average_inpatient,lowess = T,data = df_test)+
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
   theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital inpatient expenditure") +
                        ylab("Predicted risk of dying from cancer") +
                        scale_x_continuous(limits = c(0, 1)) 



cancer_outpatient <- df_test |>
                ggplot(aes(x = average_outpatient, y = Predictions_Cancer, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
  theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital outpatient expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 

cancer_primarycare <- df_test |>
                ggplot(aes(x = average_primarycare, y = Predictions_Cancer, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
  theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly primary care expenditure") +
                        ylab("Predicted risk of dying from cancer") +
                        scale_x_continuous(limits = c(0, 1)) 

cancer_prescription <- df_test |>
                ggplot(aes(x = average_prescription, y = Predictions_Cancer, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
  theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly prescription expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 


cancer_cure <- (cancer_inpatient + cancer_outpatient)/(cancer_primarycare + cancer_prescription)

pdf(".../Results/Cancer/Sex differences/cancer_cure.pdf", width = 15, height = 9)
print(cancer_cure)
dev.off()



###### CARE RELATED EXP########
#blue: non-bereaved

cancer_homecare <- df_test |>
                ggplot(aes(x = average_homecare, y = Predictions_Cancer, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
  theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly homecare expenditure") +
                        ylab("Predicted risk of dying from cancer") +
                        scale_x_continuous(limits = c(0, 1)) 

cancer_residential <- df_test |>
                ggplot(aes(x = average_residential, y = Predictions_Cancer, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
  theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly residential care expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 15)) 


cancer_care <- cancer_homecare + cancer_residential

pdf(".../Results/Cancer/Sex differences/cancer_care.pdf", width = 15, height = 5.5)
print(cancer_care)
dev.off()


```

#### CVD

```{r}
#writing recipe for the formula and making the dummy variables
xg_CVD_recipe <- recipe(surv_CVD3 ~ bereaved01 + affluence.factor +                            sex + age_bereavement.factor + 
                           children_n.factor + comorbidities_factor +
                           immigration_status + average_inpatient +
                           average_outpatient + average_prescription +
                           average_primarycare + average_residential +
                           average_homecare, data = df_train) |> 
  step_dummy(all_nominal_predictors(),one_hot = T)

#xg boost with computation engine: xgboost
xgb_CVD_spec <- boost_tree(trees = tune(),
                              mtry = tune(),
                              min_n = tune(),
                              tree_depth = tune(),
                              stop_iter = tune(),
                              learn_rate = tune()
                              ) |> 
                              set_engine("xgboost", validation = 0.2) |> 
                              set_mode("classification")

#workflow
xgb_CVD_wf <- workflow(xg_CVD_recipe, xgb_CVD_spec)


#tune hyperparameters (defines how our model is structured). Use the train df with crossvalidation.

set.seed(29)

doParallel::registerDoParallel()

xgb_CVD_rs <- tune_race_anova(xgb_CVD_wf,
                          df_train_cv,
                          grid = 30,
                          metrics = metric_set(brier_class),
                          control = control_race(verbose_elim = TRUE))

#plot results
plot_race(xgb_CVD_rs)

#Grid of hyperparameters
show_best(xgb_CVD_rs)


#finalize workflow
xgb_CVD_last <- xgb_CVD_wf |> 
    finalize_workflow(select_best(xgb_CVD_rs, 'brier_class')) |> 
    last_fit(df_split)

#metrics
xgb_CVD_last |> collect_metrics()

#predictions
xgb_CVD_last |> collect_predictions()


#group
xgb_CVD_full_preds <- xgb_CVD_last$.predictions

xgb_CVD_full_preds <- xgb_CVD_full_preds[[1]][2]

xgb_CVD_full_preds <- as.matrix(xgb_CVD_full_preds)



#Specify datadist
dd <- datadist(df_train)

options(datadist = "dd")

# evaluate performance
fit_cvd_total_score <- Score(list('LRM NULL' = fit0_cv,
                              'LRM Simple' = fit_cv_simple,
                              'LRM Full' = fit_cv_arv,
                              'Xgboost Full' = xgb_CVD_full_preds), 
                               formula = surv_CVD3 ~ 1,
                               data = df_test,
                               summary = 'ipa', plots = 'calibration',
                               metrics = c('AUC', 'brier'), 
                               se.fit = T)

plotCalibration(fit_cvd_total_score, models = "Xgboost Full", rug = F, auc.in.legend = F, brier.in.legend = F, xlab = "Predicted risk of dying from CVD", ylab = "Observed frequency of the predicted risk", col = c("#56B4E9", "#CC79A7"))



```

##### Prediction plots

```{r}
#put the predictions from the xgb into the df. 
df_test$Predictions_CVD <-  xgb_CVD_full_preds

#personal colour scheme
m_color <- c("#56B4E9", "#CC79A7", "#0072B2", "#E69F00", "#009E73")


#Differences in predicted risks for bereaved and non-bereaved.
boxplot_CVD <- df_test |>
          ggplot(aes(x = bereaved01, y = Predictions_CVD)) +
          geom_boxplot(outlier.shape = NA, color = "navy", fill = "#56B4E9", alpha=0.6) + 
          theme_minimal(base_size = 14) +
          facet_wrap(vars(sex))+
          scale_y_continuous(limits = c(0, 0.15)) + 
          xlab("Bereavement status") +
          ylab("Predicted risk of dying from CVD")


pdf(".../Results/CVD/boxplot_cvd.pdf", width = 7, height = 7)
print(boxplot_CVD)
dev.off()


#sex differenes and age
sex_age_CVD <- df_test |>
  ggplot(aes(x = age_bereavement, y = Predictions_CVD, fill=bereaved01, colour = bereaved01)) +
  geom_smooth(show.legend = F) +
  facet_wrap(vars(sex))+
  theme_minimal(base_size = 14) +
  scale_color_manual(values = m_color)+ 
  scale_fill_manual(values = m_color)+  
  xlab("Age at bereavement") +
  ylab("Predicted risk of dying from CVD") +
  scale_x_continuous(limits = c(65, 100)) 

#save
pdf(".../Results/CVD/Sex differences/sex_age_CVD.pdf", width = 7, height = 7)
print(sex_age_CVD)
dev.off()


#Differences in predicted risks for bereaved and non-bereaved. 
df_test |>
  ggplot(aes(x = bereaved01, y = Predictions_CVD)) +
  geom_boxplot(outlier.shape = NA) + 
  theme_minimal(base_size = 14) +
  facet_wrap(vars(sex))+
  scale_y_continuous(limits = c(0, 0.15)) + 
  xlab("Bereavement status") +
  ylab("Predicted risk of dying from CVD")


#distribution of risks
CVD_risk_distribution <- df_test |> 
              ggplot(aes(x = Predictions_CVD, fill = surv_CVD3)) + 
              geom_density(alpha = 0.4, show.legend = F) +
              xlab("Predicted risk of dying from CVD") +
              theme_minimal(base_size = 14) +
              scale_color_manual(values = m_color)+ 
              scale_fill_manual(values = m_color)

#save
pdf(".../Results/CVD/Sex differences/CVD_risk_distribution.pdf", width = 7, height = 7)
print(CVD_risk_distribution)
dev.off()

############ CURE RELATED EXP #################
#blue: non-bereaved

#expenditures, sex diffrences
CVD_inpatient <- df_test |>
                ggplot(aes(x = average_inpatient, y = Predictions_CVD, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend =F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital inpatient expenditure") +
                        ylab("Predicted risk of dying from CVD") +
                        scale_x_continuous(limits = c(0, 1)) 

CVD_outpatient <- df_test |>
                ggplot(aes(x = average_outpatient, y = Predictions_CVD, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital outpatient expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 

CVD_primarycare <- df_test |>
                ggplot(aes(x = average_primarycare, y = Predictions_CVD, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly primary care expenditure") +
                        ylab("Predicted risk of dying from CVD") +
                        scale_x_continuous(limits = c(0, 1)) 

CVD_prescription <- df_test |>
                ggplot(aes(x = average_prescription, y = Predictions_CVD, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly prescription expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 


CVD_cure <- (CVD_inpatient + CVD_outpatient)/(CVD_primarycare + CVD_prescription)

pdf(".../Results/CVD/Sex differences/CVD_cure.pdf", width = 15, height = 9)
print(CVD_cure)
dev.off()



###### CARE RELATED EXP########
#blue: non-bereaved

CVD_homecare <- df_test |>
                ggplot(aes(x = average_homecare, y = Predictions_CVD, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly homecare expenditure") +
                        ylab("Predicted risk of dying from CVD") +
                        scale_x_continuous(limits = c(0, 1)) 

CVD_residential <- df_test |>
                ggplot(aes(x = average_residential, y = Predictions_CVD, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly residential care expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 15)) 


CVD_care <- CVD_homecare + CVD_residential

pdf(".../Results/CVD/Sex differences/CVD_care.pdf", width = 15, height = 5.5)
print(CVD_care)
dev.off()

```

##### 

#### Dementia and Parkinsons

```{r}
#writing recipe for the formula and making the dummy variables
xg_neuro_recipe <- recipe(surv_neuro3 ~ bereaved01 + affluence.factor +                            sex + age_bereavement.factor + 
                           children_n.factor + comorbidities_factor +
                           immigration_status + average_inpatient +
                           average_outpatient + average_prescription +
                           average_primarycare + average_residential +
                           average_homecare, data = df_train) |> 
  step_dummy(all_nominal_predictors(),one_hot = T)

#xg boost with computation engine: xgboost
xgb_neuro_spec <- boost_tree(trees = tune(),
                              mtry = tune(),
                              min_n = tune(),
                              tree_depth = tune(),
                              stop_iter = tune(),
                              learn_rate = tune()
                              ) |> 
                              set_engine("xgboost", validation = 0.2) |> 
                              set_mode("classification")

#workflow
xgb_neuro_wf <- workflow(xg_neuro_recipe, xgb_neuro_spec)


#tune hyperparameters (defines how our model is structured). Use the train df with crossvalidation.

set.seed(29)

doParallel::registerDoParallel()

xgb_neuro_rs <- tune_race_anova(xgb_neuro_wf,
                          df_train_cv,
                          grid = 30,
                          metrics = metric_set(brier_class),
                          control = control_race(verbose_elim = TRUE))

#plot results
plot_race(xgb_neuro_rs)

#Grid of hyperparameters
show_best(xgb_neuro_rs)


#finalize workflow
xgb_neuro_last <- xgb_neuro_wf |> 
    finalize_workflow(select_best(xgb_neuro_rs, 'brier_class')) |> 
    last_fit(df_split)

#metrics
xgb_neuro_last |> collect_metrics()

#predictions
xgb_neuro_last |> collect_predictions()


#group
xgb_neuro_full_preds <- xgb_neuro_last$.predictions

xgb_neuro_full_preds <- xgb_neuro_full_preds[[1]][2]

xgb_neuro_full_preds <- as.matrix(xgb_neuro_full_preds)


#Specify datadist
dd <- datadist(df_train)

options(datadist = "dd")

# evaluate performance
fit_neuro_total_score <- Score(list('LRM NULL' = fit0_neuro,
                              'LRM Simple' = fit_neuro_simple,
                              'LRM Full' = fit_neuro_arv,
                              'Xgboost Full' = xgb_neuro_full_preds), 
                               formula = surv_neuro3 ~ 1,
                               data = df_test,
                               summary = 'ipa', plots = 'calibration',
                               metrics = c('AUC', 'brier'), 
                               se.fit = T)

plotCalibration(fit_neuro_total_score, models = "Xgboost Full", rug = F, auc.in.legend = F, brier.in.legend = F, xlab = "Predicted risk of dying from Dementia or Parkinson's disease", ylab = "Observed frequency of the predicted risk", col = c("#56B4E9", "#CC79A7"))
```

##### Prediction plots

```{r}
#put the predictions from the xgb into the df. 
df_test$Predictions_neuro <-  xgb_neuro_full_preds


#Differences in predicted risks for bereaved and non-bereaved.
boxplot_neuro <- df_test |>
          ggplot(aes(x = bereaved01, y = Predictions_neuro)) +
          geom_boxplot(outlier.shape = NA, color = "navy", fill = "#56B4E9", alpha=0.6) + 
          theme_minimal(base_size = 14) +
          facet_wrap(vars(sex))+
          scale_y_continuous(limits = c(0, 0.15)) + 
          xlab("Bereavement status") +
          ylab("Predicted risk of dying from dementia or Parkinson's")


pdf(".../Results/Neuro/boxplot_neuro.pdf", width = 7, height = 7)
print(boxplot_neuro)
dev.off()



#distribution of risks
neuro_risk_distribution <- df_test |> 
              ggplot(aes(x = Predictions_neuro, fill = surv_neuro3)) + 
              geom_density(alpha = 0.4, show.legend = F) +
              xlab("Predicted risk of dying from Dementia or Parkinson's") +
              theme_minimal(base_size = 14) +
              scale_color_manual(values = m_color)+ 
              scale_fill_manual(values = m_color)

#save
pdf(".../Results/Neuro/Sex differences/neuro_risk_distribution.pdf", width = 7, height = 7)
print(neuro_risk_distribution)
dev.off()

############ CURE RELATED EXP #################
#blue: non-bereaved

#expenditures, sex diffrences
neuro_inpatient <- df_test |>
                ggplot(aes(x = average_inpatient, y = Predictions_neuro, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend =F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital inpatient expenditure") +
                        ylab("Predicted risk of dying from Dementia or Parkinson's") +
                        scale_x_continuous(limits = c(0, 1)) 

neuro_outpatient <- df_test |>
                ggplot(aes(x = average_outpatient, y = Predictions_neuro, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital outpatient expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 

neuro_primarycare <- df_test |>
                ggplot(aes(x = average_primarycare, y = Predictions_neuro, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly primary care expenditure") +
                        ylab("Predicted risk of dying from Dementia or Parkinson's") +
                        scale_x_continuous(limits = c(0, 1)) 

neuro_prescription <- df_test |>
                ggplot(aes(x = average_prescription, y = Predictions_neuro, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly prescription expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 


neuro_cure <- (neuro_inpatient + neuro_outpatient)/(neuro_primarycare + neuro_prescription)

pdf(".../Results/Neuro/Sex differences/neuro_cure.pdf", width = 15, height = 9)
print(neuro_cure)
dev.off()



###### CARE RELATED EXP########
#blue: non-bereaved

neuro_homecare <- df_test |>
                ggplot(aes(x = average_homecare, y = Predictions_neuro, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly homecare expenditure") +
                        ylab("Predicted risk of dying from Dementia or Parkinson's") +
                        scale_x_continuous(limits = c(0, 1)) 

neuro_residential <- df_test |>
                ggplot(aes(x = average_residential, y = Predictions_neuro, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly residential care expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 15)) 


neuro_care <- neuro_homecare + neuro_residential

pdf(".../Results/Neuro/Sex differences/neuro_care.pdf", width = 15, height = 5.5)
print(neuro_care)
dev.off()
```

#### Diabetes

```{r}
#writing recipe for the formula and making the dummy variables
xg_dm_recipe <- recipe(surv_dm3 ~ bereaved01 + affluence.factor +                            sex + age_bereavement.factor + 
                           children_n.factor + comorbidities_factor +
                           immigration_status + average_inpatient +
                           average_outpatient + average_prescription +
                           average_primarycare + average_residential +
                           average_homecare, data = df_train) |> 
  step_dummy(all_nominal_predictors(),one_hot = T)

#xg boost with computation engine: xgboost
xgb_dm_spec <- boost_tree(trees = tune(),
                              mtry = tune(),
                              min_n = tune(),
                              tree_depth = tune(),
                              stop_iter = tune(),
                              learn_rate = tune()
                              ) |> 
                              set_engine("xgboost", validation = 0.2) |> 
                              set_mode("classification")

#workflow
xgb_dm_wf <- workflow(xg_dm_recipe, xgb_dm_spec)


#tune hyperparameters (defines how our model is structured). Use the train df with crossvalidation.

set.seed(29)

doParallel::registerDoParallel()

xgb_dm_rs <- tune_race_anova(xgb_dm_wf,
                          df_train_cv,
                          grid = 30,
                          metrics = metric_set(brier_class),
                          control = control_race(verbose_elim = TRUE))

#plot results
plot_race(xgb_dm_rs)

#Grid of hyperparameters
show_best(xgb_dm_rs)


#finalize workflow
xgb_dm_last <- xgb_dm_wf |> 
    finalize_workflow(select_best(xgb_dm_rs, 'brier_class')) |> 
    last_fit(df_split)

#metrics
xgb_dm_last |> collect_metrics()

#predictions
xgb_dm_last |> collect_predictions()

#group
xgb_dm_full_preds <- xgb_dm_last$.predictions

xgb_dm_full_preds <- xgb_dm_full_preds[[1]][2]

xgb_dm_full_preds <- as.matrix(xgb_dm_full_preds)


#Specify datadist
dd <- datadist(df_train)

options(datadist = "dd")

# evaluate performance
fit_dm_total_score <- Score(list('LRM NULL' = fit0_dm,
                              'LRM Simple' = fit_dm_simple,
                              'LRM Full' = fit_dm_arv,
                              'Xgboost Full' = xgb_dm_full_preds), 
                               formula = surv_dm3 ~ 1,
                               data = df_test,
                               summary = 'ipa', plots = 'calibration',
                               metrics = c('AUC', 'brier'), 
                               se.fit = T)

plotCalibration(fit_dm_total_score, models = "Xgboost Full", rug = F, auc.in.legend = F, brier.in.legend = F, xlab = "Predicted risk of dying from diabetes", ylab = "Observed frequency of the predicted risk", col = c("#56B4E9", "#CC79A7"), round = F)

plotCalibration(fit_dm_total_score, models = c("Xgboost Full", "LRM Full"), rug = F, auc.in.legend = F, brier.in.legend = F, xlab = "Predicted risk of dying from diabetes", ylab = "Observed frequency of the predicted risk", col = c("#56B4E9", "#CC79A7"), round = F)
```

##### Prediction plots

```{r}
#put the predictions from the xgb into the df. 
df_test$Predictions_dm <-  xgb_dm_full_preds


#Differences in predicted risks for bereaved and non-bereaved. 
boxplot_dm <- df_test |>
          ggplot(aes(x = bereaved01, y = Predictions_dm)) +
          geom_boxplot(outlier.shape = NA, color = "navy", fill = "#56B4E9", alpha=0.6) + 
          theme_minimal(base_size = 14) +
          facet_wrap(vars(sex))+
          scale_y_continuous(limits = c(0, 0.15)) + 
          xlab("Bereavement status") +
          ylab("Predicted risk of dying from diabetes")


pdf(".../Results/Diabetes/boxplot_dm.pdf", width = 7, height = 7)
print(boxplot_dm)
dev.off()


#distribution of risks
dm_risk_distribution <- df_test |> 
              ggplot(aes(x = Predictions_dm, fill = surv_dm3)) + 
              geom_density(alpha = 0.4, show.legend = F) +
              xlab("Predicted risk of dying from diabetes") +
              theme_minimal(base_size = 14) +
              scale_color_manual(values = m_color)+ 
              scale_fill_manual(values = m_color)

#save
pdf(".../Results/Diabetes/Sex differences/dm_risk_distribution.pdf", width = 7, height = 7)
print(dm_risk_distribution)
dev.off()

############ CURE RELATED EXP #################
#blue: non-bereaved

#expenditures, sex diffrences
dm_inpatient <- df_test |>
                ggplot(aes(x = average_inpatient, y = Predictions_dm, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend =F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital inpatient expenditure") +
                        ylab("Predicted risk of dying from diabetes") +
                        scale_x_continuous(limits = c(0, 1)) 

dm_outpatient <- df_test |>
                ggplot(aes(x = average_outpatient, y = Predictions_dm, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital outpatient expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 

dm_primarycare <- df_test |>
                ggplot(aes(x = average_primarycare, y = Predictions_dm, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly primary care expenditure") +
                        ylab("Predicted risk of dying from diabetes") +
                        scale_x_continuous(limits = c(0, 1)) 

dm_prescription <- df_test |>
                ggplot(aes(x = average_prescription, y = Predictions_dm, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly prescription expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 


dm_cure <- (dm_inpatient + dm_outpatient)/(dm_primarycare + dm_prescription)

pdf(".../Results/Diabetes/Sex differences/dm_cure.pdf", width = 15, height = 9)
print(dm_cure)
dev.off()



###### CARE RELATED EXP########
#blue: non-bereaved

dm_homecare <- df_test |>
                ggplot(aes(x = average_homecare, y = Predictions_dm, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly homecare expenditure") +
                        ylab("Predicted risk of dying from diabetes") +
                        scale_x_continuous(limits = c(0, 1)) 

dm_residential <- df_test |>
                ggplot(aes(x = average_residential, y = Predictions_dm, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly residential care expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 15)) 


dm_care <- dm_homecare + dm_residential

pdf(".../Results/Diabetes/Sex differences/dm_care.pdf", width = 15, height = 5.5)
print(dm_care)
dev.off()
```

#### Digestive

```{r}
#writing recipe for the formula and making the dummy variables
xg_digestive_recipe <- recipe(surv_digestive3 ~ bereaved01 + affluence.factor +                            sex + age_bereavement.factor + 
                           children_n.factor + comorbidities_factor +
                           immigration_status + average_inpatient +
                           average_outpatient + average_prescription +
                           average_primarycare + average_residential +
                           average_homecare, data = df_train) |> 
  step_dummy(all_nominal_predictors(),one_hot = T)

#xg boost with computation engine: xgboost
xgb_digestive_spec <- boost_tree(trees = tune(),
                              mtry = tune(),
                              min_n = tune(),
                              tree_depth = tune(),
                              stop_iter = tune(),
                              learn_rate = tune()
                              ) |> 
                              set_engine("xgboost", validation = 0.2) |> 
                              set_mode("classification")

#workflow
xgb_digestive_wf <- workflow(xg_digestive_recipe, xgb_digestive_spec)


#tune hyperparameters (defines how our model is structured). Use the train df with crossvalidation.

set.seed(29)

doParallel::registerDoParallel()

xgb_digestive_rs <- tune_race_anova(xgb_digestive_wf,
                          df_train_cv,
                          grid = 30,
                          metrics = metric_set(brier_class),
                          control = control_race(verbose_elim = TRUE))

#plot results
plot_race(xgb_digestive_rs)

#Grid of hyperparameters
show_best(xgb_digestive_rs)


#finalize workflow
xgb_digestive_last <- xgb_digestive_wf |> 
    finalize_workflow(select_best(xgb_digestive_rs, 'brier_class')) |> 
    last_fit(df_split)

#metrics
xgb_digestive_last |> collect_metrics()

#predictions
xgb_digestive_last |> collect_predictions()


#group
xgb_digestive_full_preds <- xgb_digestive_last$.predictions

xgb_digestive_full_preds <- xgb_digestive_full_preds[[1]][2]

xgb_digestive_full_preds <- as.matrix(xgb_digestive_full_preds)


#Specify datadist
dd <- datadist(df_train)

options(datadist = "dd")

# evaluate performance
fit_digestive_total_score <- Score(list('LRM NULL' = fit0_digestive,
                              'LRM Simple' = fit_digestive_simple,
                              'LRM Full' = fit_digestive_arv,
                              'Xgboost Full' = xgb_digestive_full_preds), 
                               formula = surv_digestive3 ~ 1,
                               data = df_test,
                               summary = 'ipa', plots = 'calibration',
                               metrics = c('AUC', 'brier'), 
                               se.fit = T)

#digestive_calibration <- plotCalibration(fit_digestive_total_score, models = "Xgboost Full", rug = F, auc.in.legend = F, brier.in.legend = F, xlab = "Predicted risk of dying from digestive diseases", ylab = "Observed frequency of the predicted risk", col = c("#56B4E9", "#CC79A7"))


```

#### Psychiatric

```{r}
#writing recipe for the formula and making the dummy variables
xg_psychiatric_recipe <- recipe(surv_psychiatric3 ~ bereaved01 + affluence.factor +                            sex + age_bereavement.factor + 
                           children_n.factor + comorbidities_factor +
                           immigration_status + average_inpatient +
                           average_outpatient + average_prescription +
                           average_primarycare + average_residential +
                           average_homecare, data = df_train) |> 
  step_dummy(all_nominal_predictors(),one_hot = T)

#xg boost with computation engine: xgboost
xgb_psychiatric_spec <- boost_tree(trees = tune(),
                              mtry = tune(),
                              min_n = tune(),
                              tree_depth = tune(),
                              stop_iter = tune(),
                              learn_rate = tune()
                              ) |> 
                              set_engine("xgboost", validation = 0.2) |> 
                              set_mode("classification")

#workflow
xgb_psychiatric_wf <- workflow(xg_psychiatric_recipe, xgb_psychiatric_spec)


#tune hyperparameters (defines how our model is structured). Use the train df with crossvalidation.

set.seed(29)

doParallel::registerDoParallel()

xgb_psychiatric_rs <- tune_race_anova(xgb_psychiatric_wf,
                          df_train_cv,
                          grid = 30,
                          metrics = metric_set(brier_class),
                          control = control_race(verbose_elim = TRUE))

#plot results
plot_race(xgb_psychiatric_rs)

#Grid of hyperparameters
show_best(xgb_psychiatric_rs)


#finalize workflow
xgb_psychiatric_last <- xgb_psychiatric_wf |> 
    finalize_workflow(select_best(xgb_psychiatric_rs, 'brier_class')) |> 
    last_fit(df_split)

#metrics
xgb_psychiatric_last |> collect_metrics()

#predictions
xgb_psychiatric_last |> collect_predictions()


#group
xgb_psychiatric_full_preds <- xgb_psychiatric_last$.predictions

xgb_psychiatric_full_preds <- xgb_psychiatric_full_preds[[1]][2]

xgb_psychiatric_full_preds <- as.matrix(xgb_psychiatric_full_preds)



#Specify datadist
dd <- datadist(df_train)

options(datadist = "dd")

# evaluate performance
fit_psychiatric_total_score <- Score(list('LRM NULL' = fit0_psychiatric,
                              'LRM Simple' = fit_psychiatric_simple,
                              'LRM Full' = fit_psychiatric_arv,
                              'Xgboost Full' = xgb_psychiatric_full_preds), 
                               formula = surv_psychiatric3 ~ 1,
                               data = df_test,
                               summary = 'ipa', plots = 'calibration',
                               metrics = c('AUC', 'brier'), 
                               se.fit = T)

psychiatric_calibration <- plotCalibration(fit_psychiatric_total_score, models = "Xgboost Full", rug = F, auc.in.legend = F, brier.in.legend = F, xlab = "Predicted risk of dying from psychiatric diseases and suicide", ylab = "Observed frequency of the predicted risk", col = c("#56B4E9", "#CC79A7"), round = F)

```

##### 

#### Respiratory

```{r}
#writing recipe for the formula and making the dummy variables
xg_respiratory_recipe <- recipe(surv_respiratory3 ~ bereaved01 + affluence.factor +                            sex + age_bereavement.factor + 
                           children_n.factor + comorbidities_factor +
                           immigration_status + average_inpatient +
                           average_outpatient + average_prescription +
                           average_primarycare + average_residential +
                           average_homecare, data = df_train) |> 
  step_dummy(all_nominal_predictors(),one_hot = T)

#xg boost with computation engine: xgboost
xgb_respiratory_spec <- boost_tree(trees = tune(),
                              mtry = tune(),
                              min_n = tune(),
                              tree_depth = tune(),
                              stop_iter = tune(),
                              learn_rate = tune()
                              ) |> 
                              set_engine("xgboost", validation = 0.2) |> 
                              set_mode("classification")

#workflow
xgb_respiratory_wf <- workflow(xg_respiratory_recipe, xgb_respiratory_spec)


#tune hyperparameters (defines how our model is structured). Use the train df with crossvalidation.

set.seed(29)

doParallel::registerDoParallel()

xgb_respiratory_rs <- tune_race_anova(xgb_respiratory_wf,
                          df_train_cv,
                          grid = 30,
                          metrics = metric_set(brier_class),
                          control = control_race(verbose_elim = TRUE))

#plot results
plot_race(xgb_respiratory_rs)

#Grid of hyperparameters
show_best(xgb_respiratory_rs)


#finalize workflow
xgb_respiratory_last <- xgb_respiratory_wf |> 
    finalize_workflow(select_best(xgb_respiratory_rs, 'brier_class')) |> 
    last_fit(df_split)

#metrics
xgb_respiratory_last |> collect_metrics()

#predictions
xgb_respiratory_last |> collect_predictions()


#group
xgb_respiratory_full_preds <- xgb_respiratory_last$.predictions

xgb_respiratory_full_preds <- xgb_respiratory_full_preds[[1]][2]

xgb_respiratory_full_preds <- as.matrix(xgb_respiratory_full_preds)


#Specify datadist
dd <- datadist(df_train)

options(datadist = "dd")

# evaluate performance
fit_respiratory_total_score <- Score(list('LRM NULL' = fit0_resp,
                              'LRM Simple' = fit_resp_simple,
                              'LRM Full' = fit_resp_arv,
                              'Xgboost Full' = xgb_respiratory_full_preds), 
                               formula = surv_respiratory3 ~ 1,
                               data = df_test,
                               summary = 'ipa', plots = 'calibration',
                               metrics = c('AUC', 'brier'), 
                               se.fit = T)

plotCalibration(fit_respiratory_total_score, models = "Xgboost Full", rug = F, auc.in.legend = F, brier.in.legend = F, xlab = "Predicted risk of dying from respiratory diseases", ylab = "Observed frequency of the predicted risk", col = c("#56B4E9", "#CC79A7"))

```

##### Prediction plots

```{r}
#put the predictions from the xgb into the df. 
df_test$Predictions_resp <-  xgb_respiratory_full_preds



#Differences in predicted risks for bereaved and non-bereaved. 
boxplot_resp <- df_test |>
          ggplot(aes(x = bereaved01, y = Predictions_resp)) +
          geom_boxplot(outlier.shape = NA, color = "navy", fill = "#56B4E9", alpha=0.6) + 
          theme_minimal(base_size = 14) +
          facet_wrap(vars(sex))+
          scale_y_continuous(limits = c(0, 0.15)) + 
          xlab("Bereavement status") +
          ylab("Predicted risk of dying from respiratory diseases")


pdf(".../Results/Respiratory/boxplot_resp.pdf", width = 7, height = 7)
print(boxplot_resp)
dev.off()



#distribution of risks
resp_risk_distribution <- df_test |> 
              ggplot(aes(x = Predictions_resp, fill = surv_respiratory3)) + 
              geom_density(alpha = 0.4, show.legend = F) +
              xlab("Predicted risk of dying from respiratory diseases") +
              theme_minimal(base_size = 14) +
              scale_color_manual(values = m_color)+ 
              scale_fill_manual(values = m_color)

#save
pdf(".../Results/Respiratory/Sex differences/resp_risk_distribution.pdf", width = 7, height = 7)
print(resp_risk_distribution)
dev.off()

############ CURE RELATED EXP #################
#blue: non-bereaved

#expenditures, sex diffrences
resp_inpatient <- df_test |>
                ggplot(aes(x = average_inpatient, y = Predictions_resp, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend =F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital inpatient expenditure") +
                        ylab("Predicted risk of dying from respiratory diseases") +
                        scale_x_continuous(limits = c(0, 1)) 

resp_outpatient <- df_test |>
                ggplot(aes(x = average_outpatient, y = Predictions_resp, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly hospital outpatient expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 

resp_primarycare <- df_test |>
                ggplot(aes(x = average_primarycare, y = Predictions_resp, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly primary care expenditure") +
                        ylab("Predicted risk of dying from respiratory diseases") +
                        scale_x_continuous(limits = c(0, 1)) 

resp_prescription <- df_test |>
                ggplot(aes(x = average_prescription, y = Predictions_resp, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly prescription expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 1)) 


resp_cure <- (resp_inpatient + resp_outpatient)/(resp_primarycare + resp_prescription)

pdf(".../Results/Respiratory/Sex differences/resp_cure.pdf", width = 15, height = 9)
print(resp_cure)
dev.off()



###### CARE RELATED EXP########
#blue: non-bereaved

resp_homecare <- df_test |>
                ggplot(aes(x = average_homecare, y = Predictions_resp, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly homecare expenditure") +
                        ylab("Predicted risk of dying from respiratory diseases") +
                        scale_x_continuous(limits = c(0, 1)) 

resp_residential <- df_test |>
                ggplot(aes(x = average_residential, y = Predictions_resp, 
                           fill=bereaved01, colour = bereaved01)) +
                        geom_smooth(show.legend = F) +
                        facet_wrap(vars(sex))+
                        theme_minimal(base_size = 14) +
                        theme(strip.text=element_text(size=14))+
                        scale_color_manual(values = m_color)+ 
                        scale_fill_manual(values = m_color)+  
                        xlab("Average weekly residential care expenditure") +
                        ylab(label = NULL) +
                        scale_x_continuous(limits = c(0, 15)) 


resp_care <- resp_homecare + resp_residential

pdf(".../Results/Respiratory/Sex differences/resp_care.pdf", width = 15, height = 5.5)
print(resp_care)
dev.off()
```

## Evaluating the AUC and brier score

Evaluating the model performance with the AUC, brier score and IPA.

### AUC

```{r}
df_AUC_cancer <- fit_cancer_total_score$AUC$score |> 
  as.data.frame() |> 
  mutate(AUC = round(AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |> 
  mutate(Cause = "Cancer")


df_AUC_cv <- fit_cvd_total_score$AUC$score |> 
  as.data.frame() |> 
  mutate(AUC = round(AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |> 
  mutate(Cause = "CV")

df_AUC_neuro <- fit_neuro_total_score$AUC$score |> 
  as.data.frame() |> 
  mutate(AUC = round(AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |> 
  mutate(Cause = "Neuro")

df_AUC_dm <- fit_dm_total_score$AUC$score |> 
  as.data.frame() |> 
  mutate(AUC = round(AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |> 
  mutate(Cause = "Diabetes")

df_AUC_digestive <- fit_digestive_total_score$AUC$score |> 
  as.data.frame() |> 
  mutate(AUC = round(AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |> 
  mutate(Cause = "Diagestive")

df_AUC_psychiatric <- fit_psychiatric_total_score$AUC$score |> 
  as.data.frame() |> 
  mutate(AUC = round(AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |> 
  mutate(Cause = "Psychiatric")

df_AUC_resp <- fit_respiratory_total_score$AUC$score |> 
  as.data.frame() |> 
  mutate(AUC = round(AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |> 
  mutate(Cause = "Respiratory")


df_AUC_together <- df_AUC_cancer |> 
  full_join(df_AUC_cv) |> 
  full_join(df_AUC_neuro) |> 
  full_join(df_AUC_dm) |> 
  full_join(df_AUC_digestive) |> 
  full_join(df_AUC_psychiatric) |> 
  full_join(df_AUC_resp) |> 
  select(Cause, model, AUC, lower, upper)


write.xlsx(df_AUC_together, ".../Results/AUC.xlsx")
```

### Delta AUC

```{r}

df_AUC_cancer <- fit_cancer_total_score$AUC$contrasts |> 
  as.data.frame() |> 
  mutate(delta.AUC = round(delta.AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |> 
  mutate(Cause = "Cancer")

df_AUC_cv <- fit_cvd_total_score$AUC$contrasts |> 
  as.data.frame() |> 
  mutate(delta.AUC = round(delta.AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "CV")

df_AUC_neuro <- fit_neuro_total_score$AUC$contrasts |> 
  as.data.frame() |> 
  mutate(delta.AUC = round(delta.AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Neuro")

df_AUC_dm <- fit_dm_total_score$AUC$contrasts |> 
  as.data.frame() |> 
  mutate(delta.AUC = round(delta.AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Diabetes")

df_AUC_digestive <- fit_digestive_total_score$AUC$contrasts |> 
  as.data.frame() |> 
  mutate(delta.AUC = round(delta.AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Diagestive")


df_AUC_psychiatric <- fit_psychiatric_total_score$AUC$contrasts |> 
  as.data.frame() |> 
  mutate(delta.AUC = round(delta.AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Psychiatric")

df_AUC_resp <- fit_respiratory_total_score$AUC$contrasts |> 
  as.data.frame() |> 
  mutate(delta.AUC = round(delta.AUC * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Respiratory")


df_delta_AUC_together <- df_AUC_cancer |> 
  full_join(df_AUC_cv) |> 
  full_join(df_AUC_neuro) |> 
  full_join(df_AUC_dm) |> 
  full_join(df_AUC_digestive) |> 
  full_join(df_AUC_psychiatric) |> 
  full_join(df_AUC_resp) |> 
  select(Cause, model, reference, delta.AUC, lower, upper)


write.xlsx(df_delta_AUC_together, ".../Results/delta_AUC.xlsx")
```

### Brier

```{r}
df_brier_cancer <- fit_cancer_total_score$Brier$score |> 
  as.data.frame() |> 
  mutate(Brier = round(Brier * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Cancer")

df_brier_cv <- fit_cvd_total_score$Brier$score |> 
  as.data.frame() |> 
  mutate(Brier = round(Brier * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "CV")

df_brier_neuro <- fit_neuro_total_score$Brier$score |> 
  as.data.frame() |> 
  mutate(Brier = round(Brier * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Neuro")

df_brier_dm <- fit_dm_total_score$Brier$score |> 
  as.data.frame() |> 
  mutate(Brier = round(Brier * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Diabetes")

df_brier_digestive <- fit_digestive_total_score$Brier$score |> 
  as.data.frame() |> 
  mutate(Brier = round(Brier * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Digestive")


df_brier_psychiatric <- fit_psychiatric_total_score$Brier$score |> 
  as.data.frame() |> 
  mutate(Brier = round(Brier * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Psychiatric")

df_brier_resp <- fit_respiratory_total_score$Brier$score |> 
  as.data.frame() |> 
  mutate(Brier = round(Brier * 100,2), 
         upper = round(upper * 100,2), 
         lower = round(lower*100,2)) |>
  mutate(Cause = "Respiratory")



df_brier_together2 <- df_brier_cancer |> 
  full_join(df_brier_cv) |> 
  full_join(df_brier_neuro) |> 
  full_join(df_brier_dm) |> 
  full_join(df_brier_digestive) |> 
  full_join(df_brier_psychiatric) |> 
  full_join(df_brier_resp) |> 
  select(Cause, model, Brier, lower, upper, IPA)


write.xlsx(df_brier_together2, ".../Results/brier_ipa.xlsx")
```

### Delta brier

```{r}
df_brier_cancer <- fit_cancer_total_score$Brier$contrasts |> 
  as.data.frame() |> 
  mutate(delta.Brier = round(delta.Brier * 100,5), 
         upper = round(upper * 100,5), 
         lower = round(lower*100,5)) |>
  mutate(Cause = "Cancer")

df_brier_cv <- fit_cvd_total_score$Brier$contrasts |> 
  as.data.frame() |> 
  mutate(delta.Brier = round(delta.Brier * 100,5), 
         upper = round(upper * 100,5), 
         lower = round(lower*100,5)) |>
  mutate(Cause = "CV")

df_brier_neuro <- fit_neuro_total_score$Brier$contrasts |> 
  as.data.frame() |> 
  mutate(delta.Brier = round(delta.Brier * 100,5), 
         upper = round(upper * 100,5), 
         lower = round(lower*100,5)) |>
  mutate(Cause = "Neuro")

df_brier_dm <- fit_dm_total_score$Brier$contrasts |> 
  as.data.frame() |> 
  mutate(delta.Brier = round(delta.Brier * 100,5), 
         upper = round(upper * 100,5), 
         lower = round(lower*100,5)) |>
  mutate(Cause = "Diabetes")

df_brier_digestive <- fit_digestive_total_score$Brier$contrasts |> 
  as.data.frame() |> 
  mutate(delta.Brier = round(delta.Brier * 100,5), 
         upper = round(upper * 100,5), 
         lower = round(lower*100,5)) |>
  mutate(Cause = "Digestive")


df_brier_psychiatric <- fit_psychiatric_total_score$Brier$contrasts |> 
  as.data.frame() |> 
  mutate(delta.Brier = round(delta.Brier * 100,5), 
         upper = round(upper * 100,5), 
         lower = round(lower*100,5)) |>
  mutate(Cause = "Psychiatric")

df_brier_resp <- fit_respiratory_total_score$Brier$contrasts |> 
  as.data.frame() |> 
  mutate(delta.Brier = round(delta.Brier * 100,5), 
         upper = round(upper * 100,5), 
         lower = round(lower*100,5)) |>
  mutate(Cause = "Respiratory")


df_brier_together <- df_brier_cancer |> 
  full_join(df_brier_cv) |> 
  full_join(df_brier_neuro) |> 
  full_join(df_brier_dm) |> 
  full_join(df_brier_digestive) |> 
  full_join(df_brier_psychiatric) |> 
  full_join(df_brier_resp) |> 
  select(Cause, model, reference, delta.Brier, lower, upper)


write.xlsx(df_brier_together, ".../Results/delta_brier.xlsx")

```
